---
title: 'Electronic supplementary material: Modelling results'
author: 'The impact of control strategies and behavioural changes on the elimination of Ebola from Lofa County, Liberia'
date: 'Sebastian Funk, Iza Ciglenecki, Amanda Tiffany, Etienne Gignoux, Anton Camacho, Rosalind M. Eggo, Adam J. Kucharski, W. John Edmunds, Josephus Bolongei, Phillip Azuma, Peter Clement, Alpha Tamba, Esther Sterk, Barbara Telfer, Gregory Engel, Lucy Anne Parker, Motoi Suzuki, Nico Heijenberg, Bruce Reeder'
output:
  pdf_document
---

```{r setup, echo = FALSE, cache = FALSE}
knitr::opts_chunk$set(cache = TRUE, fig.cap = '', fig.align = "center", message = FALSE, tidy = TRUE, warning = FALSE)
```

# Introduction

This document contains the R code necessary to reproduce the results in "Epidemiology, events and control strategies towards eliminating of Ebola from Lofa County, Liberia." All code and data used here are in a [github repository](http://github.com/sbfnk/ebola_lofa).

## Required packages
```{r packages}
cran_packages <- c("devtools","dplyr", "tidyr", "cowplot", "stringi", "truncnorm")
github_packages <- c("libbi/RBi", "sbfnk/RBi.helpers")
```

```{r install, eval = FALSE}
for (package in cran_packages)
{
    install.packages(package)
}

library('devtools')

for (package in github_packages)
{
    install_github(package)
}
```

# Generate MCMC chains

Generating MCMC chains requires a working install of [libbi](http://libbi.org/) v1.2.0.

Edit the lines starting `code_dir <- ` and `output_dir <-` in the script 'sample_posterior.r' from the R/ directory in the github repository to point to the correct directories, and run it using

```{sh eval = FALSE}
Rscript sample_posterior.r -n 10000 -p 1000 -r independent -t $NUM_THREADS -i 10 -q -l -f \
                           -o ebola_lofa_$NB -c 16384 -b $NB
```

where you should set `$NUM_THREADS` to the number of concurrent threads to run (i.e., the number of CPU cores available) and `$NB` to a running number (1,2,3,...) for each chain to generate (100 in the manuscript).

# Analyse MCMC chains

Set the `output_dir` to the value used in `sample.posterior.r`

```{r eval = FALSE}
output_dir <- 'insert your directory here'
```

```{r echo = FALSE}
output_dir <- "~/Data/Ebola/Lofa/"
```

Load required packages
```{r load}
for (package in cran_packages)
{
    library(sub("^.*/", "", package), character.only = TRUE)
}
for (package in github_packages)
{
    library(sub("^.*/", "", package), character.only = TRUE)
}
```

Get MCMC chains

```{r get_chains}
mcmc_files <- list.files(output_dir, pattern = "ebola_lofa_independent_[0-9]+\\.rds", full.names = TRUE)

res <- list()

model <- NULL
obs <- NULL
input <- NULL
data <- NULL

for (file in mcmc_files)
{
    res[[file]] <- readRDS(paste0(file))
    if (is.null(model)) model <- bi_model(sub("\\.rds$", ".bi", file))
    if (is.null(obs)) obs <- readRDS(sub("\\.rds$", "_obs.rds", file))
    if (is.null(input)) input <- readRDS(sub("\\.rds$", "_input.rds", file))
    if (is.null(data)) data <- readRDS(sub("\\.rds$", "_data.rds", file))
}
```

Combine into one chain

```{r combine}
np_translate <- NULL
combined <- lapply(names(res[[1]]), function(x) {
  z <- rbindlist(lapply(seq_along(res), function(y) {
    data.table(res[[y]][[x]])[, unique_np := paste(np, y, sep = "_")]
  }))
  if (is.null(np_translate)) {
    np_translate <-
      data.table(unique_np = unique(z$unique_np),
                 new_np = seq_along(unique(z$unique_np)) - 1)
  }
  z <- merge(z, np_translate, by = "unique_np")
  z[, np := NULL]
  z[, unique_np := NULL]
  setnames(z, "new_np", "np")
  setkey(z, np)
  z
})
```

Save

```{r save}
names(combined) <- names(res[[1]])
saveRDS(list(traces = combined, input = input, obs = obs, data = data), paste0(output_dir, "/lofa_combined_traces.rds"))
``` 

Sample observations

```{r sample}
combined$Admissions <- copy(combined$Zh)
combined$Admissions[, mean := value]
combined$Admissions[, sd := sqrt(value)]
combined$Admissions[, value := rtruncnorm(n = .N, a = 0, mean = mean, sd = sd)]
combined$Admissions[is.na(value), value := 0]
```

# Plot

```{r plot}
p <- plot_libbi(combined, model, data = data, date.origin = as.Date("2014-06-02") - 7, date.unit = "week", data.colour = "black", densities = "histogram", plot = FALSE)
```

## State estimates
```{r states}
p$states
```

## Traces
```{r traces}
p$trace
```

## Densities
```{r densities}
p$densities
```

# Analyse

## Initial number of infected
```{r Inf}
combined$p_Inf[, list(mean = mean(value),
               min.50 = quantile(value, 0.25),
               max.50 = quantile(value, 0.75),
               min.95 = quantile(value, 0.025),
               min.95 = quantile(value, 0.975))]
```

## Mean reproduction number
```{r mean_R0}
combined$p_R0[, list(mean = mean(value),
               min.50 = quantile(value, 0.25),
               max.50 = quantile(value, 0.75),
               min.95 = quantile(value, 0.025),
               min.95 = quantile(value, 0.975))]

```

## Standard deviation of reproduction number
```{r sd_R0}
combined$p_vol_R0[, list(mean = mean(value),
                  min.50 = quantile(value, 0.25),
                  max.50 = quantile(value, 0.75),
                  min.95 = quantile(value, 0.025),
                  min.95 = quantile(value, 0.975))]
```

## Early proportion seeking healthcare
```{r early_H}
elH <- combined$H[, list(early = value[1], late = value[length(value)]), by = np]
elH[, list(median = median(early),
           min.50 = quantile(early, 0.25),
           max.50 = quantile(early, 0.75),
           min.95 = quantile(early, 0.025),
           max.95 = quantile(early, 0.975))]
```

## Late proportion seeking healthcare
```{r late_H}
elH[, list(median = median(late),
           min.50 = quantile(late, 0.25),
           max.50 = quantile(late, 0.75),
           min.95 = quantile(late, 0.025),
           max.95 = quantile(late, 0.975))]
```

## Ratio between proportion seeking healthcare at last and first data point

```{r ratio_H}
elH <- combined$H[, list(early = value[1], late = value[length(value)]), by = np]
elH[, ratio := late / early]
elH[, list(median = median(ratio),
           min.50 = quantile(ratio, 0.25),
           max.50 = quantile(ratio, 0.75),
           min.95 = quantile(ratio, 0.025),
           max.95 = quantile(ratio, 0.975))]
```
